---
title: "R Markdown with LaTeX"
author: "Rob Donald"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: '2'
urlcolor: blue
---

```{r setup, include=FALSE}
# The fig.pos = 'h' tells LaTeX to put the figure where you ask for it
knitr::opts_chunk$set(echo = TRUE,fig.height=8,fig.width=12,fig.pos = 'h')
options(width=1500)
```

```{r library_setup, include=FALSE}
suppressMessages({suppressWarnings({
    library(dplyr)
    library(readr)
    library(lubridate)
    
    library(ggplot2)
    library(RobsRUtils)
    library(pander)

})})
```

```{r echo=FALSE}
theme.report <- theme_bw() + theme(plot.title = element_text(colour="black"
                                      ,size=24
                                      ,angle=0
                                      ,hjust=0.5,vjust=1.0
                                      ,face="plain")
            ,axis.text.x = element_text(colour="black"
                                      ,size=18
                                      ,angle=0
                                      ,hjust=0.5,vjust=0.6
                                      ,face="plain")
            ,axis.text.y = element_text(colour="black"
                                      ,size=12
                                      ,angle=0
                                      ,hjust=0.0,vjust=0.6
                                      ,face="plain")
            ,axis.title = element_text(colour="black"
                                      ,size=18
                                      ,angle=0
                                      ,hjust=0.5,vjust=0.6
                                      ,face="plain")
            )
```
# Introduction

This set of notes shows some examples of using \LaTeX\ with R markdown.

The idea is to give examples which help if you are reasonably comfortable
with R markdown but want *some* of the customisation of using LaTeX without
having to become a LaTeX guru \footnote{Although if you do get proficient
in this stuff you may well get the guru title}.

# What is \LaTeX?

Well if you have been 'Knit'ing to PDF inside RStudio you have been using 
LaTeX all the time. You may have already used things like \\newpage to 
tidy up a document. The \\newpage is actually a LaTeX command which is being
detected by Knitr as it processes the document.

LaTeX is used extensively in academia particularly in any subject that requires
a lot of maths notation. It is a different way of thinking about documents. You 
write your document in plain text and the various commands then format the 
document into something that looks nice. It means you can produce a PDF document
with formulas and figures and tables without ever going near Word. But it is
not for everyone. If you need a WYSIWYG editor then this is not for you.

In the past you would need to use a LaTeX editor of some kind (e.g. the cross platform open source program [TeXMaker](http://www.xm1math.net/texmaker/)). With the advent of RStudio and 
R Markdown (*which I assume you are at least partly familiar with*) you can now
write well formatted PDF documents with very little knowledge of LaTeX. 

This set of notes will show you how you can extend your documents by using
a small amount of LaTeX code in the text portion of your .Rmd R markdown 
document. 

NOTE: if you go down this route you can't flip back to HTML output. The Knitr
package will get a bit upset and probably miss out some of the LaTeX commands. It will look like it has worked, there will just be missing bits (give it a try ...)

## Some History

LaTeX is actually a typesetting programming language and was developed
by Leslie Lamport in 1983. See this link for a more detailed
description [LaTeX History](https://en.wikipedia.org/wiki/LaTeX#History).

It is a layer of code over the top of the typesetting language \TeX which
was developed by the famous computer scientist Donald Knuth in 1978. See
this link for more details [Tex History](https://en.wikipedia.org/wiki/TeX)

# Useful YouTube links

+ [LaTeX Tutorial 1 - Creating a LaTeX Document](https://www.youtube.com/watch?v=SoDv0qhyysQ)

# Font Sizes

\begin{large}
Some large text
\end{large}

\begin{normalsize}
Some normalsize text
\end{normalsize}

Compare this default R markdown line to the above 'normalsize' text.

\begin{small}
Some small text which runs over several 
lines of text and allows us to put in more information
in a smaller space. Which can be useful if the information
is very dense.
\end{small}

\begin{footnotesize}
This is footnotesize. Useful size for tables and other hard to fit in things.
\end{footnotesize}

\newpage
# References
\begin{footnotesize}
\begin{center}Footnote Size Links\end{center}
\begin{itemize}
\item{[1] \url{https://www.stata.com/manuals13/pss.pdf} \tiny{(Accessed 2019-05-02)}}
\item{[2] \url{https://en.wikipedia.org/wiki/Sample_size_determination} \tiny{(Accessed 2019-05-03)}}
\item{[3] \url{https://en.wikipedia.org/wiki/Power_(statistics)} \tiny{(Accessed 2019-05-03)}}
\item{[4] \url{https://en.wikipedia.org/wiki/Effect_size} \tiny{(Accessed 2019-05-03)}}
\item{[5] \url{https://en.wikipedia.org/wiki/Test_statistic} \tiny{(Accessed 2019-05-03)}}

\item{[6] \url{https://blog.minitab.com/blog/adventures-in-statistics-2/understanding-analysis-of-variance-anova-and-the-f-test} \tiny{(Accessed 2019-05-03)}}
\end{itemize}
\end{footnotesize}


